Supplementary Information

### 2020-06-08 doc/manuscript/MEE_manuscript.Rmd
minor update

We then used the *climwin* package in R (REF) [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590) to identify the most significant climate driver and time window for each of the three climate variable groups. 
*Climwin* searches through multiple climate variables and over a wide range of time frames to identify which are most strongly correlated to the biological variable of interest--in this case, residuals of $log[\Delta r]$, $log[BAI]$, or $log[\Delta AGB]$.
Within *climwin*, we specified a mixed effects model using species and tree identity as random effect: `residual ~ [climate] + (1 | sp) + (1 | treeID)`. 
Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of time periods, at monthly resolution, starting in the previous May and ending in August of the current year.
We note that analyzing all species together yields the most significant climate drivers across the full set of cores from each site (our goal here), whereas identification of the top climate drivers for individual species would be optimized by analyzing each species separately.
*Climwin* can check for linear and quadratic relationships, and does k-fold cross-validation in its computation of AIC to guard against over-fitting (CLIMWIN_REF). 
We specified quadratic relationships only because (i) quadratic relationships are more consistent with known biological mechanisms (see Introduction), (ii) preliminary tests revealed that quadratic fits usually had lower AIC, and when not there tended to be little difference in AIC and the curve would approximate a straight line, (iii) modeling only quadratic improves coding efficiency and speeds up the analysis process.
For each group of candidate climate variables, we move forward with the best variable over the time window identified by climwin as a candidate climate variable for the multivariate models. 
When PET came out as the best variable in both temperature and cloud cover groups, there were only two candidate variables for the GAM. 
If PET came out as the best variable in one but not both of these groups, we compared the AICs of the top variables from each group, retaining only the stronger predictor.
We checked for colinearity among the selected climate variables and removed any variable with a variance inflation factor > 10.

@@

We get different climate senstivity results with the different metrics of growth (Fig. 5?).
Preliminary review of some early results indicates that the "landscape" of climate effects over various time windows is usually(?) similar across the three metrics of growth, but that the optimal time window or even top climate variable can shift.
(**See [ISSUE #40 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/40)**).
The underlying cause of these differences would be differences in the variance structure, where growth residuals would tend to be greater (in absolute value) when growth rate is high. Thus, use of $\Delta r$ as a growth metric would tend to place proportionaly more weight on smaller $DBHs$, whereas $\Delta AGB$ would tend to place more weight on larger individuals, with $BAI$ internediate (Fig. 4). (**Christy, please check/revise**)
The degree to which climate senstivity landscapes differ across growth metrics should depend on the size structure of the data and the existence/strength of climate-DBH interactions.

>>>>

We then used the *climwin* package in R (REF) [(van de Pol et. al. 2016)](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12590) to identify the most significant climate driver and time window for each of the three climate variable groups. 
*Climwin* searches through multiple climate variables and over a wide range of time frames to identify which are most strongly correlated to the biological variable of interest--in this case, residuals of $log[\Delta r]$, $log[BAI]$, or $log[\Delta AGB]$.
Within *climwin*, we specified a mixed effects model using species and tree identity as random effect: `residual ~ [climate] + (1 | sp) + (1 | treeID)`. 
Here, for each permutation, `climate` specifies one of the climate drivers in the climate variable group, analyzed over one of all possible combinations of time periods, at monthly resolution, starting in the previous May and ending in August of the current year.
We note that analyzing all species together yields the most significant climate drivers across the full set of cores from each site (our goal here), whereas identification of the top climate drivers for individual species would be optimized by analyzing each species separately.
*Climwin* can check for linear and quadratic relationships, and does k-fold cross-validation in its computation of AIC to guard against over-fitting (CLIMWIN_REF). 
We specified quadratic relationships only because (i) quadratic relationships are more consistent with known biological mechanisms (see Introduction), (ii) preliminary tests revealed that quadratic fits usually had lower AIC, and when not there tended to be little difference in AIC and the curve would approximate a straight line, (iii) modeling only quadratic improves coding efficiency and speeds up the analysis process.
For each group of candidate climate variables, we move forward with the best variable over the time window identified by climwin as a candidate climate variable for the multivariate models. 
If PET came out as the best variable in both temperature and cloud cover groups, there were only two candidate variables for the GAM. 
If it came out as the best variable in one but not both of these groups, it was dropped from the analysis, on the logic that it was an inferior predictor to a related variable.
We checked for colinearity among the selected climate variables and removed any variable with a variance inflation factor > 10.

@@

We get different climate senstivity results with the different metrics of growth (Fig. 5?).
Preliminary review of some early results indicates that the "landscape" of climate effects over various time windows is generally similar across the three metrics of growth, but that the optimal time window or even the top climate variable in a group can shift.
(**See [ISSUE #40 in ForestGEO-climate-sensitivity ](https://github.com/EcoClimLab/ForestGEO-climate-sensitivity/issues/40)**).

The underlying cause of these differences would be differences in the variance structure, where growth residuals would tend to be greater (in absolute value) when growth rate is high. Thus, use of $\Delta r$ as a growth metric would tend to place proportionaly more weight on smaller $DBHs$, whereas $\Delta AGB$ would tend to place more weight on larger individuals, with $BAI$ internediate (Fig. 4). (**Christy, please check/revise**)
The degree to which climate senstivity landscapes differ across growth metrics should depend on the size structure of the data and the existence/strength of climate-DBH interactions.
